---
title: "VarMC Package"
output: 
  html_document:
    toc: true
    self_contained: true
    pandoc_args: ["--citeproc"]
bibliography: references.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(VaRMC)
library(tidyquant)
library(dplyr)
library(ggplot2)
```

We include the all the functions available in our package, and the for each function we briefly explain its purpose and how its incorporated with eachother to eventually obtain the VaR value.

-Stock Dedection allows us to grab a specific stock's data off the internet to then calculate its volatility measures under a 3-state HMM, the average of it, and its current returns.

-The Correlation matrix function allows us to merge the data we have from our given portfolio.

-The Covariance Matrix function allows us to turn the portfolio of our chosen stocks into a $n$x$n$ matrix.

-The Cholesky function lets us parse through the $n$x$n$ covariance matrix to end up with a lower-triangle matrix

-With the shock function, we use it to generate correlated shocks from the Cholesky factor

-We then apply these correlated shocks to the GBM through the simulated GBM function

-Then we finally compute the portfolio P&L from these simulated returns 

-Achieving the VaR value.

# Collecting stock data
Through the `stock_detection` function, lets say we want to grab the historical data of Apple,
```{r, echo=FALSE}
symbol <- "AAPL"
apple <- tidyquant::tq_get(symbol, from = "2011-01-01", to = "2026-01-01")
dplyr::glimpse(apple)
ggplot(apple, aes(x = date, y = adjusted)) +
  geom_line() +
  labs(
    title = paste(symbol, "Adjusted Close"),
    x = "Date",
    y = "Adjusted Price"
  )
```

from here on, as we can see this is the stock info of Apple during 2011 to 2026.

We can also view Apple's current volatility and also current averages of its volatilty
# ```{r, echo=FALSE, message=FALSE, warning=FALSE}
# out <- suppressMessages(
#   suppressWarnings(
#     stock_detection(apple)
#   )
# )
# 
# cat(out$current_volatility)
# cat(out$current_mean)
# ```
```{r, eval = FALSE}
stock_detection(apple)

cat(out$current_volatility)
cat(out$current_mean)
```
# Correlation Matrix and Covariance matrix 

Let's say our portfolio consists of Apple, Nvidia, Tesla, and Visa.
```{r, message=FALSE, warning=FALSE}
apple <- tq_get("AAPL", from = "2011-01-01", to = "2026-01-01")
nvidia <- tq_get("NVDA", from = "2011-01-01", to = "2026-01-01")
tesla <- tq_get("TSLA", from = "2011-01-01", to = "2026-01-01")
visa <- tq_get("V", from = "2011-01-01", to = "2026-01-01")

portfolio <- list(
  stock_detection(nvidia),
  stock_detection(apple),
  stock_detection(tesla),
  stock_detection(visa)
)

regime_vols <- sapply(portfolio, function(x) x$current_volatility)
regime_means <- sapply(portfolio, function(x) x$current_mean)
corr_mat <- correlation_matrix_of_stock_portfolio(portfolio)
cov_mat <- covariance_matrix_of_stock_portfolio(regime_vols, corr_mat)

corr_mat
cov_mat
```

# Geometric Brownian Motion Simulations
With out covariance matrix we apply Cholesky so that we can acheieve a upper-triangle matrix with respect to our portfolio
```{r, echo=FALSE}
L <- cholesky(cov_mat)

L
```
And then for the case of 10,000 simulations
``` {r, echo=FALSE}

num_sims <- 10000
shocks <- shock_gen(L, num_sims)
shocks[1:50]
```

Finally, by applying the correlated shocks into out GBM formula, we get the simulated GBM returns 

```{r, echo=FALSE, fig.width=9, fig.height=7}
n_steps <- 10000
n_plot <- 200
asset_idx <- 1
dt <- 1/252

# This makes (mu - 0.5*vol^2) = 0 inside sim_gbm
zero_drift_means <- 0.5 * regime_vols^2

returns_mat <- matrix(NA, nrow = n_plot, ncol = n_steps)

set.seed(42)
for (step in 1:n_steps) {
  daily_shocks <- shock_gen(L, n_plot)
  daily_returns <- sim_gbm(zero_drift_means, regime_vols, dt, daily_shocks)
  returns_mat[, step] <- daily_returns[asset_idx, ]
}

cum_returns <- t(apply(returns_mat, 1, cumsum))

df <- data.frame(
  step = rep(1:n_steps, each = n_plot),
  value = as.vector(cum_returns),
  sim = factor(rep(1:n_plot, times = n_steps))
)

ggplot(df, aes(x = step, y = value, group = sim, color = sim)) +
  geom_line(linewidth = 0.4) +
  scale_color_manual(values = sample(rainbow(n_plot, s = 0.85, v = 0.85))) +
  labs(
    title = "Cumulative GBM Simulated Returns",
    x = "Number of Steps",
    y = "Cumulative Log-Return"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
```

# VaR value

Finally we will be able to calculate our Var value by computing the portfolio's simulated P&L
```{r, echo=FALSE}
time_horizon <- 10
weights <- rep(1/4, 4)
conf <- 0.95
B <- 500
ci <- 0.95
sim_return <- sim_gbm(regime_means, regime_vols, time_horizon, shocks)
value <- log_to_simple_returns(sim_return)
sim_simple_returns <- t(value)
pnl <- portfolio_pnl_from_returns(sim_simple_returns, weights)
var <- compute_var(pnl, conf)
var

var_bootstrap_ci(pnl, conf, B, ci,seed = NULL)
```
This is the final value of money that we lose if we were to choose model, under 95% C.I., time horizon of 10 days, and a even 25% weight distribution in our stock portfolio.

The bootstraps also shows us the C.I. of 95% under 500 boostrapping resamples.

# Backteting
Here we do our backtesting through Kupiec's test through the same parameters used in previous examples.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
total_returns <- portfolio[[1]]$returns_data[, c("date", "return")]
for (i in 2:length(portfolio)) {
  temp <- portfolio[[i]]$returns_data[, c("date", "return")]
  total_returns <- merge(total_returns, temp, by = "date")
}
total_returns$date <- NULL
returns_matrix <- as.matrix(total_returns)

result <- backtesting(returns_matrix, weights = rep(1/4, 4),
                      conf = 0.95, window = 500,
                      time_horizon = 10, num_sims = 10000)

cat("Kupiec POF Backtest Results\n")
cat("----------------------------\n")
cat("Total Observations:    ", result$total_obs, "\n")
cat("Expected Violations:   ", result$expected_violations, "\n")
cat("Observed Violations:   ", result$violations, "\n")
cat("Expected Rate:         ", result$expected_rate, "\n")
cat("Observed Rate:         ", round(result$observed_rate, 4), "\n")
cat("LR Statistic:          ", result$LR, "\n")
```

# Comments
The backtesting strategy of using Kupiec's test is not as likely as we hoped for a big reason of that we believe is due to the fact that when we split out data into 3 states, and the Low-volatility being the most probable, that takes up a big portion of the data, and our backtesting function also works with rolling windows instead of HMM regime vols, thus it could be underestimated.

The work was evenly distributed between the two members of the team as Jeffrey focused more on the start of the project(everything before Cholesky) and Faye worked on the end(after GBM). the work in the middle was worked by ex-member Mahan. 

There majority of A.I help was used in testcases, roxygen, Rmd, ggplot, and the Backtesting portion.

# References
