---
title: "VaRMC Package"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{VaRMC Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
nocite: "@*"
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(VaRMC)
library(tidyquant)
library(dplyr)
library(ggplot2)
```

# Goal of the Package
The goal of **VaRMC** is to estimate the Value-at-Risk (VaR) of a portfolio using Monte Carlo simulation combined with regime-switching volatility estimated from Hidden Markov Models (HMMs).
This approach allows the model to capture changes in market volatility regimes when simulating asset returns. By incorporating regime-dependent volatility into the Monte Carlo framework, the package provides a flexible tool for estimating potential portfolio losses over a specified time horizon.
Such methods are commonly used in risk management to evaluate the distribution of future portfolio returns and quantify downside risk.

We include the all the functions available in our package, and then for each function we briefly explain its purpose and how its incorporated with each other to eventually obtain the VaR value.

- `stock_detection()`: retrieve historical price data for a given stock and computes key statistics required for the simulation (Specifically, it estimates the asset’s volatility using a 3-state Hidden Markov Model (HMM), along with the mean return and the historical return series)
- `correlation_matrix_of_stock_portfolio()`: construct the correlation matrix for the assets included in the portfolio
- `covariance_matrix_of_stock_portfolio()`: convert the correlation matrix and the asset volatilities into an $n$x$n$ covariance matrix, where n represents the number of assets in the portfolio
- `cholesky()`: parse through the $n$x$n$ covariance matrix to end up with a lower-triangle matrix 
- `shock_gen()`: simulate correlated shocks for Monte Carlo  
- `sim_gbm()`: GBM-based return simulation  
- `log_to_simple_returns()`: convert log-returns to simple returns  
- `portfolio_pnl_from_returns()`: compute portfolio P&L from returns and weights  
- `compute_var()`: compute VaR at a chosen confidence level  
- `var_bootstrap_ci()`: bootstrap confidence interval for VaR  
- `backtesting()`: evaluate the performance of the VaR model using historical data (implement tests such as violation counting or Kupiec-style tests)

# Collecting stock data
Through the `stock_detection` function, lets say we want to grab the historical data of Apple,
```{r, echo=TRUE}
symbol <- "AAPL"
apple <- tidyquant::tq_get(symbol, from = "2011-01-01", to = "2026-01-01")
dplyr::glimpse(apple)
ggplot(apple, aes(x = date, y = adjusted)) +
  geom_line() +
  labs(
    title = paste(symbol, "Adjusted Closing Price"),
    x = "Date",
    y = "Adjusted Price"
  )
```

from here on, as we can see this is the stock info of Apple during 2011 to 2026.

We can also view Apple's current volatility and also current averages of its volatilty
```{r, eval = FALSE}
stock_detection(apple)

cat(out$current_volatility)
cat(out$current_mean)
```
# Correlation Matrix and Covariance matrix 

Let's say our portfolio consists of Apple, Nvidia, Tesla, and Visa.
```{r, message=FALSE, warning=FALSE}
apple <- tq_get("AAPL", from = "2011-01-01", to = "2026-01-01")
nvidia <- tq_get("NVDA", from = "2011-01-01", to = "2026-01-01")
tesla <- tq_get("TSLA", from = "2011-01-01", to = "2026-01-01")
visa <- tq_get("V", from = "2011-01-01", to = "2026-01-01")

portfolio <- list(
  stock_detection(nvidia),
  stock_detection(apple),
  stock_detection(tesla),
  stock_detection(visa)
)

regime_vols <- sapply(portfolio, function(x) x$current_volatility)
regime_means <- sapply(portfolio, function(x) x$current_mean)
corr_mat <- correlation_matrix_of_stock_portfolio(portfolio)
cov_mat <- covariance_matrix_of_stock_portfolio(regime_vols, corr_mat)

corr_mat
cov_mat
```

# Geometric Brownian Motion Simulations
With out covariance matrix we apply Cholesky so that we can acheieve a upper-triangle matrix with respect to our portfolio
```{r, echo=TRUE}
L <- cholesky(cov_mat)

L
```
And then for the case of 10,000 simulations
``` {r, echo=TRUE}

num_sims <- 10000
shocks <- shock_gen(L, num_sims)
shocks[1:50]
```

Finally, by applying the correlated shocks into out GBM formula, we get the simulated GBM returns 

```{r, echo=TRUE, fig.width=9, fig.height=7}
n_steps <- 10000
n_plot <- 200
asset_idx <- 1
dt <- 1/252

# This makes (mu - 0.5*vol^2) = 0 inside sim_gbm
zero_drift_means <- 0.5 * regime_vols^2

returns_mat <- matrix(NA, nrow = n_plot, ncol = n_steps)

set.seed(42)
for (step in 1:n_steps) {
  daily_shocks <- shock_gen(L, n_plot)
  daily_returns <- sim_gbm(zero_drift_means, regime_vols, dt, daily_shocks)
  returns_mat[, step] <- daily_returns[asset_idx, ]
}

cum_returns <- t(apply(returns_mat, 1, cumsum))

df <- data.frame(
  step = rep(1:n_steps, each = n_plot),
  value = as.vector(cum_returns),
  sim = factor(rep(1:n_plot, times = n_steps))
)

ggplot(df, aes(x = step, y = value, group = sim, color = sim)) +
  geom_line(linewidth = 0.4) +
  scale_color_manual(values = sample(rainbow(n_plot, s = 0.85, v = 0.85))) +
  labs(
    title = "Cumulative GBM Simulated Returns",
    x = "Number of Steps",
    y = "Cumulative Log-Return"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
```

### Interpretation of the GBM Simulation
The wide fan of simulated paths from `sim_gbm()` reflects the uncertainty captured by the regime volatility estimated through `stock_detection()`. Because GB compounds shocks over each time step, the spread grows with the horizon. This is expected — the model is better suited for short-horizon VaR estimation (1-10 days) than long-term forecasting. An ideal result would show tighter concentration around the drift if volatility were lower or the horizon shorter.

# VaR value

Finally we will be able to calculate our Var value by computing the portfolio's simulated P&L
```{r, echo=TRUE}
time_horizon <- 10
weights <- rep(1/4, 4)
conf <- 0.95
B <- 500
ci <- 0.95
sim_return <- sim_gbm(regime_means, regime_vols, time_horizon, shocks)
value <- log_to_simple_returns(sim_return)
sim_simple_returns <- t(value)
pnl <- portfolio_pnl_from_returns(sim_simple_returns, weights)
var <- compute_var(pnl, conf)
var

var_bootstrap_ci(pnl, conf, B, ci,seed = NULL)
```
This is the final value of money that we lose if we were to choose model, under 95% C.I., time horizon of 10 days, and a even 25% weight distribution in our stock portfolio.

The bootstraps also shows us the C.I. of 95% under 500 boostrapping resamples.

# Backtesting
Here we do our backtesting through Kupiec's test through the same parameters used in previous examples.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
total_returns <- portfolio[[1]]$returns_data[, c("date", "returns_per_stock")]
for (i in 2:length(portfolio)) {
  temp <- portfolio[[i]]$returns_data[, c("date", "returns_per_stock")]
  total_returns <- merge(total_returns, temp, by = "date")
}
total_returns$date <- NULL
returns_matrix <- as.matrix(total_returns)

result <- backtesting(returns_matrix, weights = rep(1/4, 4),
                      conf = 0.95, window = 500,
                      time_horizon = 10, num_sims = 10000)

cat("Kupiec POF Backtest Results\n")
cat("----------------------------\n")
cat("Total Observations:    ", result$total_obs, "\n")
cat("Expected Violations:   ", result$expected_violations, "\n")
cat("Observed Violations:   ", result$violations, "\n")
cat("Expected Rate:         ", result$expected_rate, "\n")
cat("Observed Rate:         ", round(result$observed_rate, 4), "\n")
cat("LR Statistic:          ", result$LR, "\n")
```

# Comments
The backtesting strategy of using Kupiec's test is not as likely as we hoped for a big reason of that we believe is due to the fact that when we split out data into 3 states, and the Low-volatility being the most probable, that takes up a big portion of the data, and our backtesting function also works with rolling windows instead of HMM regime vols, thus it could be underestimated.

The work was evenly distributed between the two members of the team as Jeffrey focused more on the start of the project(everything before Cholesky) and Faye worked on the end(after GBM). the work in the middle was worked by ex-member Mahan. 

There majority of A.I help was used in testcases, roxygen, Rmd, ggplot, Backtesting portion, and parts of the explanations given during the introduction.

# References
